---
title: "Gout management in Hospital"
author:
- Jacklyn Chin
- David Fong
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
    toc_collapsed: true
toc_depth: 3
number_sections: true
theme: lumen
date: "27th August 2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warnings=FALSE, message=FALSE}
library(ggplot2)  # plotting
library(highcharter) # plotting
library(dplyr)    # data pliers
library(tidyr)    # 'tidy' up data
library(gmodels)  # contingency tables
library(mice)     # imputation
library(corrplot) # correlation plot
library(lme4)     # linear models
library(ztable)   # cross-table display
options(ztable.type="html")
library(magrittr) # pipes
library(janitor)  # clean names
library(huxtable)
```

## Read the data

81 entries

and 'clean up' the names

```{r}
d <- read.csv("SerumUricAcid.csv",
              skip = 3, # a few rows to skip at the beginning
              check.names = TRUE, # avoid duplicate names
              stringsAsFactors = FALSE
              )
d <- d[seq(from = 1, to = 81),] # 81 entries
```

### Clean names

'Clean up the names' - e.g. spaces changed to underscores '_'

```{r results='asis'}
names(d) <- make_clean_names(names(d))
names(d)
```

### Some variables are 'categorical'

...rather than ordered.

```{r}
categorical_cols <- c("gender", "primary_care_team", "grouped_units_of_admission", "hospital_campus", "hopc",
                      "rheum_input", "mtp", "ankle", "knee", "elbow", "wrist", "mcp", "tophi",
                      "nsaid", "colchicine", "prednisolone", "wean", "ia_steroid_injections",
                      "thiazide_diuretics", "loop_diuretics", "k_sparing_diuretics",
                      "b_blocker", "ac_ei", "arb", "ccb", "a_blocker", "a_b_blocker", "moxonidine",
                      "statins", "fibrates", "ezetimibe",
                      "aspirin", "clopidogrel", "htn", "t2dm", "cvd", "chol", "ckd",
                      "obesity", "clinical_asssesed_improvement", "patient_reported_improvemnet",
                      "ult_on_admission_y_n", "commenced_during", "after_discharge", "change_dose",
                      "ceased_withheld", "d_c_instructions_present",
                      "none", "infections", "diuresis", "aki", "trauma", "cognitive_impairment",
                      "acidosis", "other_rheum", "chemo", "surgery")

d[categorical_cols] <- lapply(d[categorical_cols], factor) # apply 'factor' to those columns

```

```{r}
summary(d)
```

## Serum urate vs Already on Urate Lowering Therapy 

`serum_urate_level_mmol_l` vs `ult_on_admission_y_n`

### Boxplot

```{r}
p <- ggplot(
  data = d,
  aes(
    group = ult_on_admission_y_n,
    x = ult_on_admission_y_n,
    y = serum_urate_level_mmol_l
  )
) + geom_boxplot()

p + stat_summary(fun = mean, geom = "point", shape = 23, size = 4)
```

same plot with `highcharter`

```{r}
highchart() %>%
  hc_xAxis(type = "category") %>%
  hc_add_series_list(
    data_to_boxplot(
      data = d,
      variable = serum_urate_level_mmol_l,
      group_var = ult_on_admission_y_n
    )
  ) %>%
  hc_exporting(enabled = TRUE) # allow 'save'
```

### Wilcox unpaired

(same as Mann-Whitney U)

```{r}
wilcox.test(
  serum_urate_level_mmol_l ~ ult_on_admission_y_n,
  data = d,
  exact = FALSE, paired = FALSE
)
```

## Rheumatology input

### Age

```{r}

d_notRheumPt <- d %>%
  filter(grouped_units_of_admission != 0)

p <- ggplot(
  data = d_notRheumPt,
  aes(
    group = rheum_input,
    x = rheum_input,
    y = age
  )
) + geom_boxplot()

p + stat_summary(fun = mean, geom = "point", shape = 23, size = 4)
```

### Unit of admission

```{r results='asis'}
ztable(table(d_notRheumPt$rheum_input, d_notRheumPt$grouped_units_of_admission)) %>%
  makeHeatmap(palette="Blues")
```

```{r R.options = list(width = 1000)}
CrossTable(d_notRheumPt$rheum_input, d_notRheumPt$grouped_units_of_admission,
           fisher = TRUE)
```

```{r}
p <- ggplot(
  data = d_notRheumPt,
  aes(
    group = grouped_units_of_admission,
    x = grouped_units_of_admission,
    y = rheum_input
  )
) + geom_boxplot()

p + stat_summary(fun = mean, geom = "point", shape = 23, size = 4)
```

### Gender

```{r}
p <- ggplot(
  data = d_notRheumPt,
  aes(
    group = rheum_input,
    x = rheum_input,
    y = gender
  )
) + geom_boxplot()

p + stat_summary(fun = mean, geom = "point", shape = 23, size = 4)
```

```{r results="asis"}
ztable(table(d_notRheumPt$rheum_input, d_notRheumPt$gender)) %>%
  makeHeatmap(palette="Blues")
```

```{r}
CrossTable(d_notRheumPt$rheum_input, d_notRheumPt$gender,
           fisher = TRUE)
```

### Campus

```{r results='asis'}
ztable(table(d_notRheumPt$rheum_input, d_notRheumPt$hospital_campus)) %>%
  makeHeatmap(palette="Blues")
```
```{r}
CrossTable(d_notRheumPt$rheum_input, d_notRheumPt$hospital_campus,
           fisher = TRUE)
```

### Generalized Linear Model

```{r}
rheum_consulted <- glmer(
  rheum_input ~ gender + age + ethnicity + hopc + 
    grouped_units_of_admission + duration_of_stay +
    number_of_active_joints_involved +
    mtp + knee + ankle + elbow + wrist + mcp + tophi + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted)
```

```{r}
predictors <- d_notRheumPt[, 
  c("gender", "hopc", "grouped_units_of_admission", "duration_of_stay",
    "number_of_active_joints_involved", "mtp", "knee", "ankle",
    "elbow", "wrist", "mcp", "tophi")
] %>%
  mutate_all(function(x) as.numeric(as.character(x))) %>%
  rename(c("admission_unit" = "grouped_units_of_admission", "number_of_joints" = "number_of_active_joints_involved"))
  
rheum_consult_predictor_cor <- cor(predictors, use = "complete.obs") %>% round(3)
```

```{r}
corrplot(rheum_consult_predictor_cor,
         type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```
`number_of_active_joints_involved` heavily correlated with other variables (e.g. specific active joints), and
little role in the model so far. Remove from model

```{r}
rheum_consulted2 <- glmer(
  rheum_input ~ gender + age + ethnicity + hopc + 
    grouped_units_of_admission + duration_of_stay +
    mtp + knee + ankle + elbow + wrist + mcp + tophi + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted2)
```

MTP, MCP, wrist and knee heavily correlated. Wrist has highest p-value, remove.

```{r}
rheum_consulted3 <- glmer(
  rheum_input ~ gender + age + ethnicity + hopc + 
    grouped_units_of_admission + duration_of_stay +
    mtp + knee + ankle + elbow + mcp + tophi + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted3)
```

Remove tophi

```{r}
rheum_consulted4 <- glmer(
  rheum_input ~ gender + age + ethnicity + hopc + 
    grouped_units_of_admission + duration_of_stay +
    mtp + knee + ankle + elbow + mcp + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted4)
```

Remove ankle

```{r}
rheum_consulted5 <- glmer(
  rheum_input ~ gender + age + hopc + 
    grouped_units_of_admission + duration_of_stay +
    mtp + knee + elbow + mcp + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted5)
```

Remove mcp

```{r}
rheum_consulted6 <- glmer(
  rheum_input ~ gender + age + hopc + 
    grouped_units_of_admission + duration_of_stay +
    mtp + knee + elbow + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted6)
```

Remove duration of stay

```{r}
rheum_consulted7 <- glmer(
  rheum_input ~ gender + age + hopc + 
    grouped_units_of_admission +
    mtp + knee + elbow + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted7)
```

remove of age

```{r}
rheum_consulted8 <- glmer(
  rheum_input ~ gender + hopc + 
    grouped_units_of_admission +
    mtp + knee + elbow + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted8)
```

## Secondary gout

### medications - sub-groups

```{r}
hopc <- glmer(
  hopc ~  thiazide_diuretics + loop_diuretics + k_sparing_diuretics + other_diuretics + b_blocker + ac_ei + arb + ccb + a_blocker + a_b_blocker + moxonidine + statins + fibrates + ezetimibe + aspirin + clopidogrel + (1 | hospital_campus),
  data = d,
  family = "binomial"
)

summary(hopc)
```

### medication - major groups

```{r}
hopc2 <- glmer(
  hopc ~  (x_diuretics_used > 0) + (x_anti_htn > 0) + (x_anti_chol_used > 0) + (x_anti_platelet_therapy > 0) + (1 | hospital_campus),
  data = d,
  family = "binomial"
)

summary(hopc2)
```

### conditions

```{r}
hopc3 <- glmer(
  hopc ~  htn + t2dm + cvd + chol + ckd + (1 | hospital_campus),
  data = d,
  family = "binomial"
)
# not including obesity
summary(hopc3)
```

### co-morbidities

```{r results='asis'}
n_comorbidities_hopc <- table(d$x_total_comorbidities, d$hopc)

ztable(n_comorbidities_hopc) %>%
    makeHeatmap(palette="Blues")
```

```{r}
CrossTable(d$x_total_comorbidities, d$hopc,
           fisher = TRUE)
```
```{r}
comorbidities_hopc_df <- as.data.frame(n_comorbidities_hopc) %>%
  rename(n_comorbidities = Var1, hopc = 2)
comorbidities_hopc_df
```

```{r}
comorbidities_hopc_df_wide <- pivot_wider(
  data = comorbidities_hopc_df,
  names_from = hopc,
  values_from = Freq
) %>% 
  rename(c("Primary" = "1", "Secondary" = "2")) %>%
  mutate(Proportion_Secondary = Secondary / (Primary + Secondary))
comorbidities_hopc_df_wide
```


```{r}
highchart() %>% 
  hc_chart(type = "column") %>% 
  hc_title(text = "Co-morbidities and Secondary Gout") %>%
  hc_xAxis(title = list(text = "Number of co-morbidities")) %>%
  hc_yAxis(title = list(text = "n")) %>%
  hc_plotOptions(column = list(
    dataLabels = list(enabled = FALSE),
    stacking = "normal",
    enableMouseTracking = FALSE)
  ) %>% 
  hc_series(list(name = "Primary",
                 data = comorbidities_hopc_df_wide$Primary),
            list(name = "Secondary",
                 data = comorbidities_hopc_df_wide$Secondary)
  ) %>%
  hc_exporting(enabled = TRUE) # allow 'save'
```

```{r}
highchart() %>% 
  hc_chart(type = "line") %>% 
  hc_title(text = "Co-morbidities and Secondary Gout") %>%
  hc_xAxis(title = list(text = "Number of co-morbidities")) %>%
  hc_yAxis(title = list(text = "Proportion secondary")) %>%
  hc_series(list(name = "Secondary proportion",
                 data = comorbidities_hopc_df_wide$Proportion_Secondary)) %>%
  hc_exporting(enabled = TRUE) # allow 'save'
```
```

```{r}
hopc4 <- glmer(
  hopc ~  x_total_comorbidities + (1 | hospital_campus),
  data = d,
  family = "binomial"
)
summary(hopc4)
```