---
title: "Gout management in Hospital"
author:
- Jacklyn Chin
- David Fong
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
    toc_collapsed: true
toc_depth: 3
number_sections: true
theme: lumen
date: "27th August 2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warnings=FALSE, message=FALSE}
library(ggplot2)  # plotting
library(dplyr)    # data pliers
library(gmodels)  # contingency tables
library(mice)     # imputation
library(corrplot) # correlation plot
library(lme4)     # linear models
library(ztable)   # cross-table display
options(ztable.type="html")
library(magrittr) # pipes
library(janitor)  # clean names
library(cAIC4)     # provides Akaike Information Criterion (AIC) step-wise for lme4
library(huxtable)
```

## Read the data

81 entries

and 'clean up' the names

```{r}
d <- read.csv("SerumUricAcid.csv",
              skip = 3, # a few rows to skip at the beginning
              check.names = TRUE, # avoid duplicate names
              stringsAsFactors = FALSE
              )
d <- d[seq(from = 1, to = 81),] # 81 entries
```

### Clean names

'Clean up the names' - e.g. spaces changes to underscores '_'

currently there are 'duplicate' names e.g. 'dose_3'

```{r}
names(d) <- make_clean_names(names(d))
names(d)
```

### Some variables are 'categorical'

...rather than ordered.

```{r}
d$gender <- factor(d$gender)
d$grouped_units_of_admission <- factor(d$grouped_units_of_admission)
d$hospital_campus <- factor(d$hospital_campus)
d$hopc <- factor(d$hopc)
```

## Serum urate vs Already on Urate Lowering Therapy 

`serum_urate_level_mmol_l` vs `ult_on_admission_y_n`

### Boxplot

```{r}
p <- ggplot(
  data = d,
  aes(
    group = ult_on_admission_y_n,
    x = ult_on_admission_y_n,
    y = serum_urate_level_mmol_l
  )
) + geom_boxplot()

p + stat_summary(fun = mean, geom = "point", shape = 23, size = 4)
```

### Wilcox unpaired

(same as Mann-Whitney U)

```{r}
wilcox.test(
  serum_urate_level_mmol_l ~ ult_on_admission_y_n,
  data = d,
  exact = FALSE, paired = FALSE
)
```

## Rheumatology input

### Age

```{r}

d_notRheumPt <- d %>%
  filter(grouped_units_of_admission != 0)

p <- ggplot(
  data = d_notRheumPt,
  aes(
    group = rheum_input,
    x = rheum_input,
    y = age
  )
) + geom_boxplot()

p + stat_summary(fun = mean, geom = "point", shape = 23, size = 4)
```

### Unit of admission

```{r results='asis'}
ztable(table(d_notRheumPt$rheum_input, d_notRheumPt$grouped_units_of_admission)) %>%
  makeHeatmap(palette="Blues")
```

```{r R.options = list(width = 1000)}
CrossTable(d_notRheumPt$rheum_input, d_notRheumPt$grouped_units_of_admission,
           chisq = TRUE)
```

```{r}
p <- ggplot(
  data = d_notRheumPt,
  aes(
    group = grouped_units_of_admission,
    x = grouped_units_of_admission,
    y = rheum_input
  )
) + geom_boxplot()

p + stat_summary(fun = mean, geom = "point", shape = 23, size = 4)
```

### Gender

```{r}
p <- ggplot(
  data = d_notRheumPt,
  aes(
    group = rheum_input,
    x = rheum_input,
    y = gender
  )
) + geom_boxplot()

p + stat_summary(fun = mean, geom = "point", shape = 23, size = 4)
```

```{r results="asis"}
ztable(table(d_notRheumPt$rheum_input, d_notRheumPt$gender)) %>%
  makeHeatmap(palette="Blues")
```

```{r}
CrossTable(d_notRheumPt$rheum_input, d_notRheumPt$gender,
           chisq = TRUE)
```

### Campus

```{r results='asis'}
ztable(table(d_notRheumPt$rheum_input, d_notRheumPt$hospital_campus)) %>%
  makeHeatmap(palette="Blues")
```
```{r}
CrossTable(d_notRheumPt$rheum_input, d_notRheumPt$hospital_campus,
           chisq = TRUE)
```

### Generalized Linear Model

```{r}
rheum_consulted <- glmer(
  rheum_input ~ gender + age + ethnicity + hopc + 
    grouped_units_of_admission + duration_of_stay +
    number_of_active_joints_involved +
    mtp + knee + ankle + elbow + wrist + mcp + tophi + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted)
```

```{r}
predictors <- d_notRheumPt[, 
  c("gender", "hopc", "grouped_units_of_admission", "duration_of_stay",
    "number_of_active_joints_involved", "mtp", "knee", "ankle",
    "elbow", "wrist", "mcp", "tophi")
] %>%
  mutate_all(function(x) as.numeric(as.character(x))) %>%
  rename(c("admission_unit" = "grouped_units_of_admission", "number_of_joints" = "number_of_active_joints_involved"))
  
rheum_consult_predictor_cor <- cor(predictors, use = "complete.obs") %>% round(3)
```

```{r}
corrplot(rheum_consult_predictor_cor,
         type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```
`number_of_active_joints_involved` heavily correlated with other variables (e.g. specific active joints), and
little role in the model so far. Remove from model

```{r}
rheum_consulted2 <- glmer(
  rheum_input ~ gender + age + ethnicity + hopc + 
    grouped_units_of_admission + duration_of_stay +
    mtp + knee + ankle + elbow + wrist + mcp + tophi + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted2)
```

MTP, MCP, wrist and knee heavily correlated. Wrist has highest p-value, remove.

```{r}
rheum_consulted3 <- glmer(
  rheum_input ~ gender + age + ethnicity + hopc + 
    grouped_units_of_admission + duration_of_stay +
    mtp + knee + ankle + elbow + mcp + tophi + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted3)
```

Remove tophi

```{r}
rheum_consulted4 <- glmer(
  rheum_input ~ gender + age + ethnicity + hopc + 
    grouped_units_of_admission + duration_of_stay +
    mtp + knee + ankle + elbow + mcp + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted4)
```

Remove ankle

```{r}
rheum_consulted5 <- glmer(
  rheum_input ~ gender + age + hopc + 
    grouped_units_of_admission + duration_of_stay +
    mtp + knee + elbow + mcp + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted5)
```

Remove mcp

```{r}
rheum_consulted6 <- glmer(
  rheum_input ~ gender + age + hopc + 
    grouped_units_of_admission + duration_of_stay +
    mtp + knee + elbow + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted6)
```

Remove duration of stay

```{r}
rheum_consulted7 <- glmer(
  rheum_input ~ gender + age + hopc + 
    grouped_units_of_admission +
    mtp + knee + elbow + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted7)
```

remove of age

```{r}
rheum_consulted8 <- glmer(
  rheum_input ~ gender + hopc + 
    grouped_units_of_admission +
    mtp + knee + elbow + (1 | hospital_campus),
  data = d_notRheumPt,
  family = "binomial"
)

summary(rheum_consulted8)
```

